DATA 621 - Business Analytics and Data Mining

### 2020-10-26 HW3/HW3_Final_AA.Rmd
Starting Cleanup

---
title: "DATA 621 - Business Analytics and Data Mining"
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg, Bryan Persaud,
  Jeff Shamp
date: "10/11/2020"
output:
  pdf_document: default
subtitle: 'Fall 2020 - Group 2 - Homework #3'
---

@@

```{r loadData, include=FALSE}
library(tidyverse)
library(tidymodels)
library(stacks)
library(kernlab)
library(vip)
library(summarytools)
library(corrplot)
library(knitr)
library(rsample) # model 1 libraries
library(caret)
library(e1071)

@@

df = read.csv(paste0(urlRemote, pathGithub, fileTrain))
df<- as_tibble(df)

eval = read.csv(paste0(urlRemote, pathGithub, fileTest))
```

@@

For each predictors, we computed and drew kernel density estimate to understand their distribution. The following plots show how predictors are distributed between areas where crime rate is higher than the median, i.e. blue, and areas where crime rate is below the median, i.e. red. It is of interest to understand variables that highlight large variations between the two groups.

@@

```{r density}
# density plots
variable = names(df)
df.new = df
denplot = function(i){
  df.new$x = df.new[,variable[i]]
  ggplot(df.new, aes(x, fill = factor(target))) + 
    ggtitle(sprintf("%s", variable[i])) + 
    geom_density(alpha = 0.4, show.legend = FALSE) + 
    theme(legend.position = "none",
          plot.title = element_text(size = 10, hjust = 0.5),
          axis.title.y = element_blank(),
          axis.title.x = element_blank())
}
gridExtra::grid.arrange(denplot(1),denplot(2),denplot(3),denplot(4),
                        denplot(5),denplot(6),denplot(7),denplot(8),
                        denplot(9),denplot(10),denplot(11),denplot(12),
                        layout_matrix = rbind(c(0,1,2), c(3,4,5), c(6,7,8), c(9,10,11)))
```

@@

>>>>

---
title: "DATA 621 - Business Analytics and Data Mining"
subtitle: 'Fall 2020 - Group 2 - Homework #3'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "10/11/2020"
output:
  pdf_document: default

---

@@

```{r loadData, include=FALSE}
library(tidyverse)
library(tidymodels)
library(stacks)
library(kernlab)
library(vip)
library(summarytools)
library(corrplot)
library(knitr)
library(rsample) # model 1 libraries
library(caret)
library(e1071)
library(geoR)

@@

df <- read.csv(paste0(urlRemote, pathGithub, fileTrain))
df <- as_tibble(df)
eval <- read.csv(paste0(urlRemote, pathGithub, fileTest))
```

@@

For each predictors, we computed kernel density estimators to understand their
distribution. The following plots show how predictors are distributed between
areas where the crime rate is higher than the median (blue) and areas where the
crime rate is below the median (red). It is of interest to understand variables
that highlight large variations between the two groups.

```{r density}
# density plots
var_names <- cbind(names(df), 'factor(target)')
plot_list <- list()
for (i in seq_along(names(df))) {
   p <- ggplot(df, aes_string(x=var_names[i, 1], fill=var_names[i, 2])) +
                 ggtitle(var_names[i, 1]) +
                 geom_density(alpha = 0.4, show.legend = FALSE) + 
                 theme(legend.position = "none",
                       plot.title = element_text(size = 10, hjust = 0.5),
                       axis.title.y = element_blank(),
                       axis.title.x = element_blank())
   plot_list[i] <- p
}
gridExtra::grid.arrange(plot_list[1], plot_list[2], plot_list[3], plot_list[4],
                        plot_list[5], plot_list[6], plot_list[7], plot_list[8],
                        plot_list[9], plot_list[10], plot_list[11],
                        plot_list[12],
                        layout_matrix = rbind(c(0,1,2), c(3,4,5), c(6,7,8),
                                              c(9,10,11)))
```

@@

@@

The variable `chas` will be removed due to having statistically insignificant correlation with almost all the variables in the dataset.

```{r}
new_df <- dplyr::select(df, -chas)
```

The lower outliers at the lower 1.5 IQR will be capped at the 5th percentile while the higher outliers at the upper 1.5 IQR will be capped at the 95th percentile.

```{r}
id <- c(1:12)
for (value in id) {
  quantiles <- quantile(new_df[,value], probs = c(.25, .75), na.rm = TRUE)
  caps <- quantile(new_df[,value], probs = c(.05, .95), na.rm = T)
  H <- 1.5 * IQR(new_df[,value], na.rm = T)
  new_df[,value][new_df[,value] < (quantiles[1] - H)] <- caps[1]
  new_df[,value][new_df[,value] > (quantiles[2] + H)] <- caps[2]
}
```

We will use Boxcox for variables that are skewed.

```{r}
boxcoxfit(new_df$age)
boxcoxfit(new_df$lstat)
```

The boxcox suggest a transformation of 1.318 for age and 0.233 for lstat.

```{r}
new_df$age <- new_df$age ^ 1.318
new_df$lstat <- new_df$lstat ^ 0.233
```

```{r}
boxcoxfit(new_df$rm)
```

The boxcox suggests a transformation of 0.204 for rm.

```{r}
new_df$rm <- new_df$rm ^ 0.204
```

Plot to see the affects of the transformations.

```{r}
hist_age <- ggplot(new_df, aes(age)) + geom_histogram()
hist_lstat <- ggplot(new_df, aes(lstat)) + geom_histogram()
hist_rm <- ggplot(new_df, aes(rm)) + geom_histogram()
grid.arrange(hist_age, hist_lstat, hist_rm)
```
